---
title: "HTE simulations"
author: "Rachael Caelie (Rocky) Aikens"
date: "2/25/2020"
output: pdf_document
---



```{r setup, warning=FALSE, message = FALSE, include = FALSE}
knitr::opts_chunk$set(cache=TRUE, warning = FALSE, message = FALSE, echo = FALSE, fig.align = "center", fig.height = 4)
require(dplyr)
require(ggplot2)
require(gridExtra)
require(ggpubr)
require(knitr)
theme_set(theme_light())
source("../code/basic_sim_functions.R")
set.seed(123)
```


# Set Up With HTE

We compare the performance of propensity score matching, Mahalanobis distance matching, and pilot matching (described in the manuscript) on simulated data, varying the dimensionality of the problem, the fixed treatment to control ratio during matching, and the correlation between the true propensity and prognostic score. The generative model for all of our simulations is the following:
\begin{align*}
    X_i &\sim_{iid} \text{Normal}(0,I_p),\\
    T_i &\sim_{iid} \text{Bernoulli}\left(\frac{1}{1+\exp(-\phi(X_i))}\right),\\
    Y_i &=\tau_i T_i + \Psi(X_i) + \epsilon_i,\\
    \epsilon_i &\sim_{iid} N(0,\sigma^2),
\end{align*}
where the true propensity and prognositic scores are given by the linear combinations
\begin{align*}
    \phi(X_i) &= X_{i1}/3-c,\\
    \Psi(X_i) &=\rho X_{i1} + \sqrt{(1-\rho^2)}X_{i2},
\end{align*}
so that $\text{Cor}(\phi(X_i), \Psi(X_i)) \propto \rho$.  The constant, $c$, in the propensity score formula was chosen such that there were approximately 100 treated observations in each dataset. We consider $p=10$, $\rho = 0, 0.1,\hdots, 0.9, 1.0,$ and $k=1,\hdots, 10$. Eachmike   simulation consisted of a dataset of size $n=2000$ and was repeated $N=1000$ times.


In this set-up, the treatment effect, $\tau_i$ always has mean 1, but it is allowed to be heterogeneous between individuals. The noise in the outcome is fixed as $\sigma=1$.


For a given matching, we estimate SATT and design sensitivity $\tilde\Gamma$ using the permutation $t$-statistic from the package `sensitivtymv`.

# Treatment effect depends on X1

In this batch of simulations, we let $\tau_i = 1 + 0.25X_{i1}$.  This means that, in the whole sample, $\tau$ is normally distributed with mean 1 and standard deviation 0.25.  Since $\tau$ depends on $X_{i1}$, this models the likely scenario that individuals are selecting into the treatment group based on some prior suggestion that the treatment will be especially beneficial to them (i.e. people who are more-likely to have a high treatment effect are also more likely to be treated.)

```{r}
read_HTE_data_1000 <- function(i, path_to_file){
  filename <- paste(path_to_file, "HTE_results_",i,"_10_1000", sep = "")
  dat <- read.csv(filename) %>%
    mutate(rho = i/10)
  return(dat)
}


dat <- lapply(1:10, function(x) read_HTE_data_1000(x, "../data/HTE_sims/HTE_tau_X1/")) %>% bind_rows
```


```{r}
dat <- dat %>%
  filter( k <= 5) %>%
  mutate(squared_err = (estimate-SATT)**2,
         k = as.factor(k))

plt_data <- dat %>% 
  group_by(method, k, rho) %>% 
  summarize(Bias = mean(estimate - SATT), 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Pilot"))
```


```{r}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Pilot")))
```

```{r, fig.width=8.5, fig.height=10.5}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) +
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
  xlab(expression(paste("Correlation, ", rho)))+
  ylim(0, NA) + 
  ylab("Standard Deviation")+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method)+
  ylim(0, NA) + 
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) + 
  ylim(0, NA) + 
  scale_color_brewer(palette="RdYlBu")


ggarrange(a, b, c, d, ncol = 1, nrow = 4, common.legend = TRUE, legend = "right", labels = "AUTO")

```

Everything is more or less the same.