---
title: "Lasso Results"
output: pdf_document
---

```{r setup, warning=FALSE, message = FALSE, include = FALSE}
knitr::opts_chunk$set(cache=FALSE, warning = FALSE, message = FALSE, echo = FALSE, fig.align = "center", fig.height = 4)
require(ggplot2)
require(dplyr)
require(gridExtra)
require(ggpubr)
theme_set(theme_light())
source("../code/basic_sim_functions.R")
```

# Primary simulations from paper

For reference, the performance of each method under this formulation is given below.  

```{r}
read_data_1000 <- function(i, path_to_file){
  filename <- paste(path_to_file, "angle_sigma1_results_",i,"_10_1000", sep = "")
  dat <- read.csv(filename) %>%
    mutate(rho = i/10)
  return(dat)
}

dat <- lapply(1:10, function(x) read_data_1000(x, "../data/tuning/mu_x1_over_3_minus_3/nsim_1000/")) %>% bind_rows %>% filter(k < 6)
```

```{r}
true_tau <- 1

dat <- mutate(dat, 
              squared_err = (estimate-true_tau)**2,
              k = as.factor(k))

plt_data <- dat %>% 
  group_by(method, k, rho) %>% 
  summarize(Bias = abs(mean(estimate) - true_tau), 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Buffalo"))
```

```{r, echo = FALSE}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Buffalo")))
```

```{r, fig.width=8.5, fig.height=9.5, echo = FALSE}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) +
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
  xlab(expression(paste("Correlation, ", rho)))+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method)+
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) + 
  scale_color_brewer(palette="RdYlBu")

ggarrange(a, b, c, d, ncol = 1, common.legend = TRUE, legend = "right")
```

\pagebreak

# Lasso

Same idea, except now the score models are both fit with a lasso model.

Note that the coefficient of X1 in the propensity score model was increased (from 1/3 to 1/2) because otherwise it was common for the lasso regression to over-regularize and conclude that all coefficients were zero.

```{r}
read_data_1000_lasso <- function(i, path_to_file){
  filename <- paste(path_to_file, "lasso_results_",i,"_10_1000", sep = "")
  dat <- read.csv(filename) %>%
    mutate(rho = i/10)
  return(dat)
}
dat <- lapply(1:10, function(x) read_data_1000_lasso(x, "../data/lasso/")) %>% bind_rows
```

```{r}
true_tau <- 1

dat <- mutate(dat, 
              squared_err = (estimate-true_tau)**2,
              k = as.factor(k))

plt_data <- dat %>% 
  group_by(method, k, rho) %>% 
  summarize(Bias = abs(mean(estimate) - true_tau), 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Buffalo"))
```

```{r, echo = FALSE}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Buffalo")))
```

```{r, fig.width=8.5, fig.height=9.5, echo = FALSE}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) +
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
  xlab(expression(paste("Correlation, ", rho)))+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method)+
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) + 
  scale_color_brewer(palette="RdYlBu")

ggarrange(a, b, c, d, ncol = 1, nrow = 4, common.legend = TRUE, legend = "right")
```
