---
title: "Supplementary Figures"
subtitle: 'Pilot Designs: Using Abundant Data Thoughtfully'
output: pdf_document
---

```{r setup, warning=FALSE, message = FALSE, include = FALSE}
knitr::opts_chunk$set(cache=FALSE, warning = FALSE, message = FALSE, echo = FALSE, fig.align = "center", fig.height = 4)
require(ggplot2)
require(dplyr)
require(gridExtra)
require(ggpubr)
theme_set(theme_light())
source("../code/basic_sim_functions.R")

set.seed(123)
```

```{r}
read_data_1000_lasso <- function(i, path_to_file){
  filename <- paste(path_to_file, "lasso_results_",i,"_10_1000", sep = "")
  dat <- read.csv(filename) %>%
    mutate(rho = i/10)
  return(dat)
}
dat <- lapply(1:10, function(x) read_data_1000_lasso(x, "../data/lasso/")) %>% bind_rows
```

```{r}
true_tau <- 1

dat <- dat %>% mutate(
              squared_err = (estimate-true_tau)**2,
              k = as.factor(k))

plt_data <- dat %>%
  group_by(method, k, rho) %>% 
  summarize(Bias = mean(estimate) - true_tau, 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Pilot"))
```

```{r, echo = FALSE}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Pilot")))
```

```{r, fig.width=8.5, fig.height=8.9, echo = FALSE}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) +
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
    ylab("Standard Deviation")+
  xlab(expression(paste("Correlation, ", rho)))+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method)+
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) + 
  scale_color_brewer(palette="RdYlBu")

ggarrange(a, b, c, d, ncol = 1, nrow = 4, common.legend = TRUE, legend = "right", labels = "AUTO")
```

**Supplementary Figure 1:** Bias, standard deviation, MSE and median gamma when the score models were fit with a lasso.  All simulation parameters are the same as described in Section 4.2, except that $\phi(X_i) = X_{i1}/2 - 3$, giving more structure to the propensity score so that it could reasonably be fit with a sparsifying model.  The regularization penalty for each score model was determined using 10-fold cross validation.  In rare cases, the propensity score model with minimum cross-validation error amounted to estimating all non-intercept model coefficients as zero.  When this occurred, a very mild ($\lambda = 0.005$) regularization was used as a default, rather than the cross-validation optimum.

\pagebreak


```{r}
read_data_1000 <- function(i, path_to_file){
  filename <- paste(path_to_file, "angle_sigma1_results_",i,"_50_1000", sep = "")
  dat <- read.csv(filename) %>%
    mutate(rho = i/10)
  return(dat)
}
```

```{r}
dat <- lapply(1:10, function(x) read_data_1000(x, "../data/p_50/")) %>% bind_rows %>%
  filter(k <= 5)
```

```{r}
true_tau <- 1

dat <- dat %>% mutate(
              squared_err = (estimate-true_tau)**2,
              k = as.factor(k))

plt_data <- dat %>%
  group_by(method, k, rho) %>% 
  summarize(Bias = mean(estimate) - true_tau, 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Pilot"))
```

```{r, echo = FALSE}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Pilot")))
```

```{r, fig.width=8.5, fig.height=9.4, echo = FALSE}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) +
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
    ylab("Standard Deviation")+
  xlab(expression(paste("Correlation, ", rho)))+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method)+
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) + 
  scale_color_brewer(palette="RdYlBu")

ggarrange(a, b, c, d, ncol = 1, nrow = 4, common.legend = TRUE, legend = "right", labels = "AUTO")
```

**Supplementary Figure 2:** Bias, standard deviation, MSE, and median $\Gamma$ of matching estimators when the number of uninformative covariates is increased. All simulation parameters are the same as described in Section 4.2, except that the number of covariates, $p$, is increased to 50.

\pagebreak

```{r}
read_xSITA_data_1000 <- function(i, path_to_file, nu){
  filename <- paste(path_to_file, "xSITA_results_",i,"_10_1000_nu_", nu * 10, sep = "")
  dat <- read.csv(filename) %>%
    mutate(rho = i/10)
  return(dat)
}

dat <- lapply(1:10, function(x) read_xSITA_data_1000(x, "../data/SITA_violations/nu_2/", 0.2)) %>% bind_rows
```


```{r}
true_tau <- 1

dat <-  dat %>%
  filter(k <= 5) %>%
  mutate(squared_err = (estimate-true_tau)**2,
        k = as.factor(k))

plt_data <- dat %>% 
  group_by(method, k, rho) %>% 
  summarize(Bias = mean(estimate) - true_tau, 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Pilot"))
```


```{r}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Pilot")))
```

```{r, fig.width=8.5, fig.height=9.4, echo = FALSE}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) +
  ylim(c(0, NA)) +
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
  xlab(expression(paste("Correlation, ", rho)))+
  ylab("Standard Deviation")+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method)+
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) + 
  scale_color_brewer(palette="RdYlBu")

ggarrange(a, b, c, d, ncol = 1, nrow = 4, common.legend = TRUE, legend = "right", labels = "AUTO")

```

**Supplementary Figure 3:** Bias, standard deviation, MSE, and median $\Gamma$ of matching estimators when an unobserved confounder is present.


\pagebreak

```{r}
read_data_1000 <- function(i, path_to_file){
  filename <- paste(path_to_file, "angle_sigma1_results_",i,"_10_1000", sep = "")
  dat <- read.csv(filename) %>%
    mutate(rho = i/10)
  return(dat)
}
```

```{r}
dat <- lapply(1:10, function(x) read_data_1000(x, "../data/fewer_controls/mu_x1_over_3_minus_275/")) %>% bind_rows %>% filter (k <= 5)
```

```{r}
true_tau <- 1

dat <- mutate(dat, 
              squared_err = (estimate-true_tau)**2,
              k = as.factor(k))

plt_data <- dat %>% 
  group_by(method, k, rho) %>% 
  summarize(Bias = mean(estimate) - true_tau, 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Pilot"))
```

```{r, echo = FALSE}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Pilot")))
```

```{r, fig.width=8.5, fig.height=9.4, echo = FALSE}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) +
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
  xlab(expression(paste("Correlation, ", rho)))+
    ylab("Standard Deviation")+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method)+
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) + 
  scale_color_brewer(palette="RdYlBu")

ggarrange(a, b, c, d, ncol = 1, nrow = 4, common.legend = TRUE, legend = "right")
```

**Supplementary Figure 4:** Mahalanobis, propensity, and prognostic pilot matching performance in terms of Bias (A), Variance (B), and MSE(C) for 1:k optimal matching with a smaller control reserve ($n = 1600$)

\pagebreak

```{r}
FM_plot <- function(data, rho, title = ""){
  plt_data <- data %>%
    mutate(prog = rho*X1 + sqrt(1-rho^2)*X2, 
           prop = mu,
           t = as.factor(abs(1-t)),
           a = ifelse(t == 1, 0.9, 1)) %>%
    dplyr::select(c(t, prog, prop, a))
  
  plt <- ggplot(data = plt_data, aes( x = prop, y = prog, group = t, color = t)) + 
    geom_point(size = 1, aes(alpha = a)) +
    scale_color_brewer(palette="Set1") +
    theme(legend.position = "none", aspect.ratio=1, plot.title = element_text(hjust = 0.5, size = 12))+
    ggtitle(title) +
    ylab(expression(paste("Prognostic Score, ", Psi, "(x)", sep = ""))) +
    xlab("Propensity Score, logit(e(x))")
  
  return(plt)
}

overlap_histogram <- function(data, rho = 0){
    plt_data <- data %>% 
      mutate(prog = rho*X1 + sqrt(1-rho^2)*X2, 
           prop = mu,
           t = as.factor(abs(1-t))) %>%
      select(c(t, prog, prop))
    
    ggplot(plt_data, aes(x = prop, y = ..density.., fill = t)) +
      scale_fill_brewer(palette = "Set1", name = "", labels = c("Treated", "Control")) +
      geom_histogram(alpha = 0.4, position = "identity") +
      ylab("Density") + 
      xlab("Logit Propensity Score, logit(e(x))")
}
```

```{r, fig.height=3, fig.width=8}
a <- FM_plot(generate_data(true_mu = "X1-10/3"), rho = 0, title = "")
b <- FM_plot(generate_data(true_mu = "X1-10/3"), rho = 0.5, title = "")
c <- FM_plot(generate_data(true_mu = "X1-10/3"), rho = 0.9, title = "")
ggarrange(a,b,c, labels ="AUTO", ncol = 3)
```

&nbsp;

**Supplementary Figure 5:** Example Fisher-Mill plots for a scenario with diminished covariate overlap between treated and control individuals.  Blue dots represent control individuals, red dots represent treated individuals, and dotted lines connect matched pairs.  The correllation $\rho$ between $\phi$ and $\Psi$ is 0, 0.5, and 0.9, respectively in figures (A), (B) and (C). Simulations were carried out as described in Section 4.2, but with $\phi(X_i) = X_{i1} - 10/3$. This kept the sample size and number of treated individuals constant while increasing the separation of treated and control individuals in terms of the covariate X1.

&nbsp;
&nbsp;
&nbsp;
&nbsp;
&nbsp;
&nbsp;
&nbsp;



```{r fig.height=3, fig.width=5}
set.seed(150)
overlap_histogram(generate_data(true_mu = "X1-10/3"))
```

**Supplementary Figure 6:** Example overlayed density plot of logit propensity score in treated and control individuals for a simulation with diminished covariate overlap (in this simulation, $\rho = 0$).  Note that there were approximately 19 control observations for every treated observation for each data set, so in practice the counts for the control observations are much larger.

\pagebreak

```{r}
dat <- lapply(1:10, function(x) read_data_1000(x, "../data/mu_x1_minus_10_3rds/nsim_1000/")) %>%
  bind_rows() %>%
  filter(k <=5)
```

```{r}
true_tau <- 1

dat <- dat %>% mutate(
              squared_err = (estimate-true_tau)**2,
              k = as.factor(k))

plt_data <- dat %>%
  group_by(method, k, rho) %>% 
  summarize(Bias = mean(estimate) - true_tau, 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Pilot"))
```

```{r, echo = FALSE}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Pilot")))
```

```{r, fig.width=8.5, fig.height=9.4, echo = FALSE}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) +
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
    ylab("Standard Deviation")+
  xlab(expression(paste("Correlation, ", rho)))+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method)+
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) + 
  scale_color_brewer(palette="RdYlBu")

ggarrange(a, b, c, d, ncol = 1, nrow = 4, common.legend = TRUE, legend = "right", labels = "AUTO")
```

**Supplementary Figure 7:** Bias, standard deviation, MSE, and median $\Gamma$ from matching estimators for simulations with worse covariate overlap between treated and control observations.


\pagebreak

```{r}
dat <- lapply(1:10, function(x) read_data_1000(x, "../data/tuning/sigma_2/nsim_1000/")) %>%
  bind_rows() %>%
  filter(k <= 5)
```

```{r}
true_tau <- 1

dat <- dat %>% mutate(
              squared_err = (estimate-true_tau)**2,
              k = as.factor(k))

plt_data <- dat %>%
  group_by(method, k, rho) %>% 
  summarize(Bias = mean(estimate) - true_tau, 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Pilot"))
```

```{r, echo = FALSE}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Pilot")))
```

```{r, fig.width=8.5, fig.height=9.4, echo = FALSE}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) +
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
    ylab("Standard Deviation")+
  xlab(expression(paste("Correlation, ", rho)))+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method)+
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) + 
  scale_color_brewer(palette="RdYlBu")

ggarrange(a, b, c, d, ncol = 1, nrow = 4, common.legend = TRUE, legend = "right", labels = "AUTO")
```

**Supplementary Figure 8:** Bias, standard deviation, MSE and median gamma when the random noise contributing to the outcome is increased.  All simulation parameters are the same as described in Section 4.2, except that $\sigma = 2$ This increases the difficulty of fitting the prognostic score, diminishing the relative performance of pilot matching in terms of MSE and sensitivity.

\pagebreak

```{r}
read_HTE_data_1000 <- function(i, path_to_file){
  filename <- paste(path_to_file, "HTE_results_",i,"_10_1000", sep = "")
  dat <- read.csv(filename) %>%
    mutate(rho = i/10)
  return(dat)
}


dat <- lapply(1:10, function(x) read_HTE_data_1000(x, "../data/HTE_sims/HTE_tau_X1/")) %>% bind_rows
```


```{r}
dat <- dat %>%
  filter( k <= 5) %>%
  mutate(squared_err = (estimate-SATT)**2,
         k = as.factor(k))

plt_data <- dat %>% 
  group_by(method, k, rho) %>% 
  summarize(Bias = mean(estimate - SATT), 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Pilot"))
```


```{r}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Pilot")))
```

```{r, fig.width=8.5, fig.height=9.4}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) +
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
  xlab(expression(paste("Correlation, ", rho)))+
  ylab("Standard Deviation")+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method)+
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab(expression(paste("Correlation, ", rho)))+
  facet_wrap(~method) + 
  scale_color_brewer(palette="RdYlBu")


ggarrange(a, b, c, d, ncol = 1, nrow = 4, common.legend = TRUE, legend = "right", labels = "AUTO")

```

**Supplementary Figure 9:** Bias, standard deviation, MSE, and median $\Gamma$ from matching estimators for simulations with a heterogeneous treatment effect. Simulations were carried out as described in Section 4.2, but with the individual treatment effect determined by the covariates: $\tau(X_i) = 1 + X{i1}/4$.
