---
title: "Main Figures"
author: "Rachael Caelie (Rocky) Aikens"
date: "5/3/2019"
output: pdf_document
---


```{r setup, warning=FALSE, message = FALSE, include = FALSE}
knitr::opts_chunk$set(cache=TRUE, warning = FALSE, message = FALSE, echo = FALSE, fig.align = "center", fig.height = 4)
require(dplyr)
require(ggplot2)
require(gridExtra)
require(ggpubr)
require(knitr)
theme_set(theme_light())
source("../code/basic_sim_functions.R")
set.seed(123)

install.packages("stratamatch")
library("stratamatch")
```

# Motivations

How can we more efficiently use data?  Increasing sample size reduces standard deviation (increases precision) at approximately a rate of 1/n.  

```{r, fig.width = 3, fig.height = 2}
n_data <- data.frame(n = seq(1,50))

ggplot(n_data, aes(x = n, y = 1/sqrt(n))) + geom_line() 
```

# Set Up 

We compare the performance of propensity score matching, Mahalanobis distance matching, and pilot matching (described in the manuscript) on simulated data, varying the dimensionality of the problem, the fixed treatment to control ratio during matching, and the correlation between the true propensity and prognostic score. The generative model for all of our simulations is the following:
\begin{align*}
    X_i &\sim_{iid} \text{Normal}(0,I_p),\\
    T_i &\sim_{iid} \text{Bernoulli}\left(\frac{1}{1+\exp(-\phi(X_i))}\right),\\
    Y_i &=\tau T_i + \Psi(X_i) + \epsilon_i,\\
    \epsilon_i &\sim_{iid} N(0,\sigma^2),
\end{align*}
where the true propensity and prognositic scores are given by the linear combinations
\begin{align*}
    \phi(X_i) &= X_{i1}/3-c,\\
    \Psi(X_i) &=\rho X_{i1} + \sqrt{(1-\rho^2)}X_{i2},
\end{align*}
so that $\text{Cor}(\phi(X_i), \Psi(X_i)) \propto \rho$.  The constant, $c$, in the propensity score formula was chosen such that there were approximately 100 treated observations in each dataset. We consider $p=10$, $\rho = 0, 0.1,\hdots, 0.9, 1.0,$ and $k=1,\hdots, 10$. Each simulation consisted of a dataset of size $n=2000$ and was repeated $N=1000$ times.
We fix the treatment effect to be constant with $\tau=1$ and the noise to be $\sigma=1$.
For a given matching, we estimate ATT and design sensitivity $\tilde\Gamma$ using the permutation $t$-statistic from the package `sensitivtymv`.

\pagebreak

# Motivation

The figure below gives a heuristic representation of what this algorithm is attempting to do.  We imagine a scenario in which each individual in our data set, both treated and control, is represented in a reduced space of only two covariates: The variation determining the treatment assignment ($\phi(X_i)$), and the variation determining the outcome ($\Psi(X_i)$). These are the two features which are directly relevant to our matching: $\phi(X_i)$ balance (propensity balance) reduces bias, and $\Psi(X_i)$ balance (prognostic balance) reduces bias as well as variance and sensitivity to unobserved confounding. 

In our simulation, $\phi(X_i)$ and $\Psi(X_i)$ are known linear combinations of the covariates (see set up), so we can visualize them directly.  For simplicity, we assume that the prognosis and treatment assignment are entirely uncorrellated ($\rho = 0$), although this need not always be the case (See supplementary figures 1 and 2). Optimal mahalanobis distance matching (Figure 1A), pairs individuals who are closest in the full covariate space.  However, since only $X_{i1}$ and $X_{i2}$ are important for prognosis and treatment assignment, individuals who are close in the full covariate space may be very distant in the feature space of $\phi(X_i)$ and $\Psi(X_i)$. Propensity score matching (Figure 1B) pairs individuals who are close in the axis important for treatment assigment, $\phi(X_i)$, but not for prognosis, $\Psi(X_i)$.  This matching will reduce bias compared to the unmatched dataset, but will lose the protection from variance and unobserved confounding conferred by prognostic balance.  In contrast, if we match jointly on $\phi(X_i)$ and $\Psi(X_i)$, we obtain individuals who are close together in the feature space below.  This optimizes for both desirable types of covariate balance: prognostic and propensity.

```{r}
FM_plot <- function(data, rho){
  plt_data <- data %>%
    mutate(prog = rho*X1 + sqrt(1-rho^2)*X2, 
           prop = mu,
           t = as.factor(abs(1-t)),
           a = ifelse(t == 1, 0.9, 1)) %>%
    dplyr::select(c(t, prog, prop, a))
  
  plt <- ggplot(data = plt_data, aes( x = prop, y = prog, group = t, color = t)) + 
    geom_point(size = 1, aes(alpha = a)) +
    scale_color_brewer(palette="Set1") +
    theme(legend.position = "none", aspect.ratio=1, plot.title = element_text(hjust = 0.5, size = 12))+
    ylab(expression(paste("Prognostic Score, ", Psi, "(x)", sep = ""))) +
    xlab("Propensity Score, logit(e(x))")
  
  return(plt)
}

match_viz <- function(data, match, rho, k = 1, title = "Matching"){
  plt_data <- data %>% 
    mutate(m = match) %>%
    mutate(a = ifelse (is.na(m), 0.9, 1)) %>% 
    mutate(prog = rho*X1 + sqrt(1-rho^2)*X2, 
           prop = mu,
           t = as.factor(abs(1-t))) %>%
    dplyr::select(c(t, prog, prop, m, a))
  
  m_data <- plt_data %>% 
    filter(!is.na(m)) %>%
    arrange(m, desc(t)) %>% 
    mutate(id = rep(1:(k + 1), sum(data$t))) %>%
    dplyr::select(-c(t, a)) %>%
    group_by(m) %>%
    summarize(prop1 = first(prop), prop2 = last(prop),
              prog1 = first(prog), prog2 = last(prog)) %>%
    dplyr::select(prog1, prog2, prop1, prop2)
  
  plt <- ggplot(data = plt_data, aes( x = prop, y = prog, group = t, color = t)) + 
    geom_point(aes(alpha = a), size = 1)+
    scale_color_brewer(palette="Set1") +
    geom_segment(data = m_data, 
                 aes(x = prop1, y = prog1,
                     xend = prop2, yend = prog2),
                 color =  "black", group = NA, linetype = "dashed") +
    ggtitle( title)+
    theme(legend.position = "none", aspect.ratio=1, plot.title = element_text(hjust = 0.5, size = 12))+
    ylab(expression(paste("Prognostic Score, ", Psi, "(x)", sep = ""))) +
    xlab("Propensity Score, logit(e(x))")
  
  return(plt)
}

# like prognostic match except returns data frame and match assignments, not just the
# reformatted dataframe of outcomes by match assignment
prognostic_match_assignment <- function(df, propensity, match_assignment, prog_model, n_control) {
  df$m <- match_assignment
  df$row <- 1:nrow(df)
  n_t<- sum(df$t)

  selected <- df %>% 
    filter(!is.na(m)) %>%
    filter(t==0) %>%
    group_by(m) %>%
    sample_n(size = 1)
  
  prognostic <- lm(y ~ . - mu - t - row - m, data = selected)
  not_selected <- df[-selected$row, ]
  not_selected <- not_selected %>% 
			mutate(progscore = predict(prognostic, not_selected)) %>%
			mutate(propscore = predict(propensity, not_selected))
  prog_dist <- match_on(t ~ progscore + propscore, data = not_selected)
  prog_match <- pairmatch(prog_dist, controls = n_control, data = not_selected) 
  return(list(df = not_selected, match = prog_match, k = n_control))
}
```

```{r}
rho <- 0.5
#simulate data
df <- generate_data(N = 2000, p = 10, true_mu = "X1/3-3", rho = rho, sigma = 1)
k = 1
prop_model = formula(t ~ . - mu - y)
prog_model = formula(y ~ . - mu - t)

# mahalanobis match
mahal_dist <- match_on(prop_model, method = "mahalanobis", data = df)
m_match <- pairmatch(mahal_dist, controls = k, df)
```


```{r}
#Calculate true propensity and prognostic score, and match on the true score
oracle_df <- df %>% 
    mutate(prog = rho*X1 + sqrt(1-rho^2)*X2, 
           prop = 1/(1+exp(-(mu))))

oracle_prop_match <- pairmatch(t ~ prop, controls = k, oracle_df)
oracle_prog_match <- pairmatch(t ~ prog + prop, controls = k, oracle_df)
```

```{r, fig.width=9, fig.height= 3}
a <- FM_plot(df, rho)
df2 <- generate_data(N = 2000, p = 10, true_mu = "1.5*X1-4", rho = rho, sigma = 1)
b <- FM_plot(df2, 0.9)
df3 <- generate_data(N = 2000, p = 10, true_mu = "4*X1-5", rho = rho, sigma = 1)
c <- FM_plot(df3, 0)

ggarrange(a,b,c, ncol= 3)
```

```{r}
a <- match_viz(df, m_match, rho, title = "Mahalanobis Distance")
b <- match_viz(df, oracle_prop_match, rho, title = "Propensity")
c <- match_viz(df, oracle_prog_match, rho, title = "Propensity x Prognosis")

ggarrange(a,b,c, ncol= 3)
```

```{r}
pdf("../figures/Poster_Figure1.pdf",  width=9, height=3)
ggarrange(a,b,c, ncol= 3)
dev.off()
```

# 5 Results 

## 5.1 Pilot matching performance with a large control reserve

Below, we use the simulation parameters described in the set up to estimate the bias, variance, mse, and median gamma sensitivity of effect estimates produced from Mahalanobis distance matching, propensity score matching, and pilot matching.

```{r}
read_data_1000 <- function(i, path_to_file){
  filename <- paste(path_to_file, "angle_sigma1_results_",i,"_10_1000", sep = "")
  dat <- read.csv(filename) %>%
    mutate(rho = i/10)
  return(dat)
}

dat <- lapply(1:10, function(x) read_data_1000(x, "../data/tuning/mu_x1_over_3_minus_3/nsim_1000/")) %>% bind_rows
```

```{r}
true_tau <- 1

dat <- mutate(dat, 
              squared_err = (estimate-true_tau)**2,
              k = as.factor(k))

plt_data <- dat %>% 
  group_by(method, k, rho) %>% 
  summarize(Bias = mean(estimate) - true_tau, 
            median_gamma = median(gamma), 
            Standard.Deviation = sd(estimate),
            MSE = Bias^2 + Standard.Deviation^2) %>%
  ungroup() %>%
  mutate(method = recode(method, propensity = "Propensity", 
                         mahalanobis = "Mahalanobis", 
                         prognostic = "Propensity x Prognosis"))
```


```{r}
plt_data <- plt_data %>%
  mutate(method = factor(method, levels = c("Mahalanobis", "Propensity", "Propensity x Prognosis")))
```

```{r, fig.width=7.5, fig.height=4.5}
a <- ggplot(plt_data, aes(x = rho, y = Bias, group = k, color = k)) +
  geom_line() + geom_point() +
  xlab("Correlation of treatment and prognosis")+
  facet_wrap(~method) +
  theme(strip.text.x = element_text(size = 12, face = "bold"))+
  theme(legend.text = element_text(angle = -90, hjust = 0.4))+
  scale_color_brewer(palette="RdYlBu")

b <- ggplot(plt_data, aes(x = rho, y = Standard.Deviation, group = k, color = k)) +
  geom_line() + geom_point() + facet_wrap(~method) +
  xlab("Correlation of treatment and prognosis")+
  theme(strip.text.x = element_text(size = 12, face = "bold"))+
  ylab("Standard Deviation")+
    theme(legend.text = element_text(angle = -90, hjust = 0.4))+
  scale_color_brewer(palette="RdYlBu")

c <- ggplot(plt_data, aes(x = rho, y = MSE, group = k, color = k)) +
  geom_line() + geom_point() + 
  xlab("Correlation of treatment and prognosis")+
  facet_wrap(~method)+
  theme(strip.text.x = element_text(size = 12, face = "bold"))+
  scale_color_brewer(palette="RdYlBu")

d <- ggplot(plt_data, aes(x = rho, y = median_gamma, group = k, color = k)) +
  geom_line() + geom_point() +
  ylab(expression(paste("Median ", Gamma))) +
  xlab("Correlation of treatment and prognosis")+
  facet_wrap(~method) + 
  theme(strip.text.x = element_text(size = 12, face = "bold"))+
  scale_color_brewer(palette="RdYlBu")

ggarrange(a, b, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right")

```



```{r fig.width=8.5, fig.height=6, echo = FALSE}
p <- ggarrange(a, b, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right")

ggsave("../figures/TT_Figure2.pdf", p,  width=7.5, height=4.5)
```


```{r fig.width=8.5, fig.height=6, echo = FALSE}
ggarrange(c, d, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right")
```

```{r fig.width=8.5, fig.height=6, echo = FALSE}
p <- ggarrange(c, d, ncol = 1, nrow = 2, common.legend = TRUE, legend = "right")

ggsave("../figures/TT_Figure3.pdf", p,  width=7.5, height=4.5)
```



## 5.2 Methodological Considerations

In the previous section, we illustrated a use case in which pilot matching is very useful: There is an abundance of control individuals which overlap fairly well with the treated population, and the underlying processes dictating propensity and prognosis are easily fit with standard linear models.  In this section, we consider four design considerations which are important to the selection of the method: 

(1) Correlation of treatment and prognosis (confoundedness)

(2) Tradeoffs in sample size

(3) Tradeoffs in match quality

(4) Fitting the propensity and prognostic models


### 5.2.4 Fitting propensity and prognostic models


```{r}
# Build scores empirically for propensity and prognostic match

# build propensity score
propensity <- glm(prop_model, family = binomial(), data = df)
  
prop_match <- pairmatch(propensity, controls = k, df)
  
# 1:2 mahalanobis matching to select data to use for prognostic model
mahal_match <- pairmatch(mahal_dist, controls = 2, df) 
  
buff_match_assignment <- prognostic_match_assignment(df, propensity, mahal_match, prog_model, k)
```

```{r, fig.width=6.85, fig.height= 2.5}
a <- match_viz(df, m_match, rho, title = "Mahalanobis\nDistance")
b <- match_viz(df, prop_match, rho, title = "Estimated\nPropensity")
c <- match_viz(buff_match_assignment$df, buff_match_assignment$match, rho, title = "Estimated\nPropensity x Prognosis")

ggarrange(a,b,c, ncol= 3)
```

```{r}
pdf("../figures/Poster_Figure6.pdf",  width=6.85, height=2.5)
ggarrange(a,b,c, ncol= 3)
dev.off()
```